---
title: "Data wrangling Project"
author: "Liliana Iotebatu"
date: "2025-04-09"
output: html_document
---


# Data downloaded from: 

https://qfish.fisheries.qld.gov.au/query/charter-catch-and-effort/table#

# Research Question

What is the relationship between the number of fishing licenses issued and 
the number of days spent at sea?


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/User/OneDrive - James Cook University/Desktop/2025 subjects/Techniques 1/Data science in R/MB5370_Module_04/MB5370_Module_04/Module04_Liliana")

```

# Packages
```{r, warning=FALSE,message=FALSE} 
library(tidyverse)
library(ggplot2)
library(janitor)
```

# Load data 

```{r, warning=FALSE,message=FALSE}
Data1<-read_csv("data/Data1.csv")
```

# Data exploration and cleaning

```{r, warning=FALSE}
# Convert the first row of Data1 into header names and clean them up
Data_clean<- clean_names(
  row_to_names(Data1, row_number = 1)
)
```

```{r, warning=FALSE}
# Create a new data frame 'Combination'
Combination <- Data_clean |> 
  select(na | contains("combination")) |> 
  row_to_names(row_number = 1) |>
  mutate(across(is.character, as.numeric)) |>
  filter(!is.na(CalendarYear))
```


```{r}
# Filter out rows with NA values in the specified columns
Combination_clean <- Combination %>% filter(!is.na(Licences), !is.na(Days))
head(Combination_clean)
```


# Plot graph

```{r, warning=FALSE}
ggplot(Combination_clean, aes(x = CalendarYear, y = Licences)) +
  geom_line(color = "black", size = 1) +  # draws the line
  geom_point(color = "black", size = 1) +  # adds points for each observation
  labs(
    title = "Number of Licences by Year",
    x = "Year",
    y = "Number of Licences"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.border = element_rect(fill = NA, color = "grey", size = 1)
  )

```


**Comment**
Licences fluctuate significantly from 2000 to 2010 (with notable peaks around the late 2000s), then settle 
into lower, more stable numbers in the mid- to late 2010s, ending with a slight upward trend in recent years.


```{r, warning=FALSE, message=FALSE}
ggplot(Combination_clean, aes(x = Licences, y = Days)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "loess", se = TRUE, color = "red") +
  labs(
    title = "Relationship between fishing licences and days at sea",
    x = "no. Licences",
    y = "no. Days at Sea",
    caption = "Source: Your Data"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    panel.border = element_rect(color = "grey", fill = NA, size = 1)
  )

```


**Comment**
The plot reveals a clear positive trend between the number of licences and days spent at sea. Initially, increases in 
licences correspond to more days at sea, but the growth eventually tapers off, indicating that beyond a certain point, 
additional licences yield diminishing increases in fishing effort.








