---
title: "Data wrangling Project"
author: "Liliana Iotebatu"
date: "2025-04-09"
output: html_document
---


# Data downloaded from: 

https://qfish.fisheries.qld.gov.au/query/charter-catch-and-effort/table#

# Research Question

What is the relationship between the number of Line fishing licenses issued and 
the number of days spent at sea?


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/User/OneDrive - James Cook University/Desktop/2025 subjects/Techniques 1/Data science in R/MB5370_Module_04/MB5370_Module_04/Module04_Liliana")

```

# Packages
```{r, warning=FALSE,message=FALSE} 
library(tidyverse)
library(ggplot2)
library(janitor)
```

# Load data 

```{r, warning=FALSE,message=FALSE}
Data1<-read_csv("data/Data1.csv")
```

# Data exploration and cleaning

```{r, warning=FALSE}
# Convert the first row of Data1 into header names and clean them up
Data_clean<- clean_names(
  row_to_names(Data1, row_number = 1)
)
```


```{r, warning=FALSE,message=FALSE}
# Create new dataframe for Line
Line <- Data_clean |> 
  select(na | contains("line")) |> 
  row_to_names(row_number = 1) |>
  mutate(across(is.character, as.numeric)) |>
  filter(!is.na(CalendarYear))
```


```{r, warning=FALSE,message=FALSE}
# Filter out rows with NA values in the specified columns
Line_clean <- Line %>% filter(!is.na(Licences), !is.na(Days))
head(Line_clean)
```


# Plot graph

```{r, warning=FALSE, message=FALSE}
plot1<- ggplot(Line_clean, aes(x = factor(CalendarYear), y = Licences)) +
  geom_col(fill = "steelblue") +       # Use geom_col() to create bars
  coord_flip() +                       # Flip coordinates for a horizontal layout
  labs(
    title = "Total no. of Licences by Year",
    x = "Year",
    y = "Number of Licenses"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.border = element_rect(fill = NA, color = "grey", size = 1)
  )
```


**Comment**
This bar chart shows yearly licence totals, revealing notable peaks around the mid-2000s 
and lower counts in other years. The variation suggests shifting licensing activity over time, 
with some standout spikes against more moderate levels in other periods.

```{r, warning=FALSE, message=FALSE}
plot2<- ggplot(Line_clean, aes(x = Licences, y = Days)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "loess", se = TRUE, color = "red") +
  labs(
    title = "Relationship between no. fishing licences and annual days at sea",
    x = "no. Licences",
    y = "no. Days at Sea (annual total)",
    
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 11),
    panel.border = element_rect(color = "grey", fill = NA, size = 1)
  )
```


**Comment**
This graph shows a strong positive relationship between the number of line fishing licences and annual days at sea.
The red LOESS curve, along with its confidence band, indicates that as licences increase, fishing effort(measured by days at sea) generally rises, despite some variation across the dataset.

# Save the plot

```{r, warning=FALSE, message=FALSE}
ggsave("plot1.jpeg", plot = plot1, width = 6, height = 4, dpi = 300, units = "in")
ggsave("plot2.jpeg", plot = plot2, width = 6, height = 4, dpi = 300, units = "in")

```






